The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-27 manuscript.Rmd
fix cross refs

\CRANpkg{drake} has been demonstrated to run inside containers for high reproducibility\footnote{See for example \href{https://github.com/joelnitta/pleurosoriopsis}{https://github.com/joelnitta/pleurosoriopsis} or \href{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}, the latter even running in a CI, see~\nameref{cicd}.}, but also how to use \CRANpkg{future} package's function `makeClusterPSOCK()` to send parts of the workflow to a Docker image for execution\footnote{\href{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}} \citep[see package's function documentation;~][]{future_2020}.

@@

\citet{gentleman_statistical_2007} coined the term **Research Compendium** for a dynamic document together with supporting data and code.
They use the R package system \cite{core_writing_1999} as for the functional prototype to structuring, validation, and distribution of research compendia.
This concept has been taken up and extended\footnote{See full literature list at \href{https://research-compendium.science/}{https://research-compendium.science/}.}, not the least by applying containerisation and other methods for managing computing environments---see Section&nbsp;\nameref{envs}.
Containers give the researcher an isolated environment to assemble these research pipelines with specific versions of software.
Research workflows in containers are safe from contamination from other activities occuring on the researcher's computer, for example the installation of newest version of packages for teaching demonstrations.
Given the users in this scenario, i.e., often academics with limited formal software development training, templates and assistance with containers around research compendia is essential.
In many fields we see that a typical unit of research for a container is a research report or journal article, where the container holds the compendium, or self-contained set of data (or connections to data elsewhere) and code files needed to fully reproduce the article \citep{marwick_packaging_2018}.
The package \pkg{rrtools} ([https://github.com/benmarwick/rrtools](https://github.com/benmarwick/rrtools)) provides a template and convienence functions to apply good practices for research compendia, including a starter `Dockerfile`.
Images of compendium containers can be hosted on services such as Docker&nbsp;Hub for convienient sharing among collaborators and others. 
Similarly, packages such as \pkg{containerit} and \pkg{dockerfiler} can be used to manage the `Dockerfile` to be archived with a compendium on a data repository (e.g. [Zenodo](https://zenodo.org/), [Dataverse](https://dataverse.org/), [Figshare](https://figshare.com/), [OSF](https://osf.io/)).
A typial compendium's `Dockerfile` will pull a rocker image fixed to a specific version of R, and install R packages from the MRAN repository to ensure the package versions are tied to a specific date, rather than the most recent version.
A more extreme case is the _dynverse_ project ([https://dynverse.org/](https://dynverse.org/)), which packages over 50 computational methods with different environments (R, Python, C++, etc.) in Docker images, which can be executed from R.
_dynverse_ uses a CI (cf.&nbsp;\nameref{cicd}) to build Rocker-derived containers, test them, and push them to Docker&nbsp;Hub if the tests succeed.

>>>>

\CRANpkg{drake} has been demonstrated to run inside containers for high reproducibility\footnote{See for example \href{https://github.com/joelnitta/pleurosoriopsis}{https://github.com/joelnitta/pleurosoriopsis} or \href{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}, the latter even running in a continuous integration platform (cf.~\nameref{development}.}, but also how to use \CRANpkg{future} package's function `makeClusterPSOCK()` to send parts of the workflow to a Docker image for execution\footnote{\href{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}} \citep[see package's function documentation;~][]{future_2020}.

@@

\citet{gentleman_statistical_2007} coined the term **Research Compendium** for a dynamic document together with supporting data and code.
They use the R package system \cite{core_writing_1999} as for the functional prototype to structuring, validation, and distribution of research compendia.
This concept has been taken up and extended\footnote{See full literature list at \href{https://research-compendium.science/}{https://research-compendium.science/}.}, not the least by applying containerisation and other methods for managing computing environments---see Section&nbsp;\nameref{envs}.
Containers give the researcher an isolated environment to assemble these research pipelines with specific versions of software.
Research workflows in containers are safe from contamination from other activities occuring on the researcher's computer, for example the installation of newest version of packages for teaching demonstrations.
Given the users in this scenario, i.e., often academics with limited formal software development training, templates and assistance with containers around research compendia is essential.
In many fields we see that a typical unit of research for a container is a research report or journal article, where the container holds the compendium, or self-contained set of data (or connections to data elsewhere) and code files needed to fully reproduce the article \citep{marwick_packaging_2018}.
The package \pkg{rrtools} ([https://github.com/benmarwick/rrtools](https://github.com/benmarwick/rrtools)) provides a template and convienence functions to apply good practices for research compendia, including a starter `Dockerfile`.
Images of compendium containers can be hosted on services such as Docker&nbsp;Hub for convienient sharing among collaborators and others. 
Similarly, packages such as \pkg{containerit} and \pkg{dockerfiler} can be used to manage the `Dockerfile` to be archived with a compendium on a data repository (e.g. [Zenodo](https://zenodo.org/), [Dataverse](https://dataverse.org/), [Figshare](https://figshare.com/), [OSF](https://osf.io/)).
A typial compendium's `Dockerfile` will pull a rocker image fixed to a specific version of R, and install R packages from the MRAN repository to ensure the package versions are tied to a specific date, rather than the most recent version.
A more extreme case is the _dynverse_ project ([https://dynverse.org/](https://dynverse.org/)), which packages over 50 computational methods with different environments (R, Python, C++, etc.) in Docker images, which can be executed from R.
_dynverse_ uses a continuous integreation (CI) platform to build Rocker-derived images, test them, and publish them on Docker&nbsp;Hub if the tests succeed.
